---
title: "IST707 Team Project: HR attrition prediction"
author: "Bomin Powers, Karen Kim"
date: "12/18/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
# Load the libraries
library(psych)
library(Hmisc)
library(lubridate)
library(factoextra)
library(e1071)
library(rpart)
library(maptree)
library(rattle)
library(car)
library(caret)
library(arules)
library(arulesViz)
library(ggplot2)
```


# import the data
```{r}
hrdata<- read.csv("D:/Documents/school/IST707/project/HRdata.csv")
summary(hrdata)
```

# data cleansing
```{r}
# remove unnecessary columns
hrdata<- subset(hrdata, select = -c(X,Date_of_termination))
```

# convert character data type to nominal types to conduct correlation analysis

```{r}
# convert age column(int) to nominal type

# check how age's data distributed
hist(hrdata$Age, breaks = 15, main = paste("Histogram of Age"))
# split age variable into somewhat equal-sized groups
hrdata$ageG <- as.factor(cut2(hrdata$Age, g=6))
# check how each group range look like
summary(hrdata$ageG)
## group1 = 18:28, group2 = 29:32, group3 = 33:36, group4 = 37:40, group5 = 41:46, group6 = 47:60
# now, convert age column(int) to nominal type
hrdata$ageG <- as.numeric(cut2(hrdata$Age, g=6))
hrdata<- subset(hrdata, select = -c(Age))
```

```{r}
# convert Attrition column(cha) to nominal type
# convert Yes to 1 and No to 0
hrdata$Attrition <- ifelse(hrdata$Attrition=="Yes",1,0)
```


```{r}
# convert BusinessTravel column(cha) to nominal type
hrdata$BusinessTravel = as.numeric(as.factor(hrdata$BusinessTravel))
# Non-Travel == 1, Travel_Frequently == 2,Travel_Rarely ==3

# convert Department column(cha) to nominal type
hrdata$Department = as.numeric(as.factor(hrdata$Department))
# Human Resources == 1, Sales == 3,Research & Development ==2

# convert JobRole column(cha) to nominal type
hrdata$JobRole = as.numeric(as.factor(hrdata$JobRole))
# Healthcare Representative == 1, Human Resources == 2, Laboratory Technician ==3, 
# Manager=4, Manufacturing Director==5, Research Director ==6,
# Research Scientist ==7, Sales Executive==8, Sales Representative==9

# convert MaritalStatus column(cha) to nominal type
hrdata$MaritalStatus = as.numeric(as.factor(hrdata$MaritalStatus))
# Divorced == 1, Married == 2,Single == 3

# convert Higher_Education column(cha) to nominal type
hrdata$Higher_Education = as.numeric(as.factor(hrdata$Higher_Education))
# 12th == 1, Graduation == 2,PHD == 3, Post-Graduation == 4

# convert Status_of_leaving column(cha) to nominal type
hrdata$Status_of_leaving = as.numeric(as.factor(hrdata$Status_of_leaving))
# Better Opportunity == 1,

# convert Source_of_Hire column(cha) to nominal type
hrdata$Source_of_Hire = as.numeric(as.factor(hrdata$Source_of_Hire))
# Job Event == 1, Portal == 2,Recruiter == 3, Walk-in == 4

# convert Job_mode column(cha) to nominal type
hrdata$Job_mode = as.numeric(as.factor(hrdata$Job_mode))
# Contract == 1, FullTime == 2, Part Time == 3
```


```{r}
# convert DistanceFromHome column(int) to nominal type

# check how DistanceFromHome's data distributed
hist(hrdata$DistanceFromHome, breaks = 15, main = paste("Histogram of DistanceFromHome"))
# split DistanceFromHome variable into somewhat equal-sized groups
hrdata$DistanceFromHomeG <- as.factor(cut2(hrdata$DistanceFromHome, g=8))
# check how each group range look like
summary(hrdata$DistanceFromHomeG)
# group1 = 1, group2 = 2, group3 =3:4, group4 = 5:7, group5 = 8:9, group6 = 10:14, group7 = 15:22, group8=23:29
# now, convert DistanceFromHome column(int) to nominal type
hrdata$DistanceFromHomeG <- as.numeric(cut2(hrdata$DistanceFromHome, g=8))
hrdata<- subset(hrdata, select = -c(DistanceFromHome))
```

```{r}
# convert Gender column(cha) to binary type

# convert Male to 1 and Female to 0
hrdata$Gender <- ifelse(hrdata$Gender =="Male",1,0)
```


```{r}
# convert MonthlyIncome column(int) to nominal type

# check how MonthlyIncome's data distributed
hist(hrdata$MonthlyIncome, breaks = 15, main = paste("Histogram of MonthlyIncome"))
# split MonthlyIncome variable into somewhat equal-sized groups
hrdata$MonthlyIncomeG <- as.factor(cut2(hrdata$MonthlyIncome, g=6))
# check how each group range look like
summary(hrdata$MonthlyIncomeG)
# group1 = 1009:2560, group2 = 2561:3632, group3 =3633:4929, group4 = 4930:6537, group5 = 6538:10551, group6 = 10552:19999
# now, convert age column(int) to nominal type
hrdata$MonthlyIncomeG <- as.numeric(cut2(hrdata$MonthlyIncome, g=6))
hrdata<- subset(hrdata, select = -c(MonthlyIncome))
```

```{r}
# check how NumCompaniesWorked's data distributed
hist(hrdata$NumCompaniesWorked, breaks = 15, main = paste("Histogram of NumCompaniesWorked"))
## this could be a problem since the values are clustered between 0 to 2 years but, for now keep this way.
summary(as.factor(hrdata$NumCompaniesWorked))
```

```{r}
# convert OverTime column(cha) to binary type

# convert Yes to 1 and No to 0
hrdata$OverTime <- ifelse(hrdata$OverTime =="Yes",1,0)
```

```{r}
# convert PercentSalaryHike column(int) to nominal type

# check how PercentSalaryHike's data distributed
hist(hrdata$PercentSalaryHike, breaks = 15, main = paste("Histogram of PercentSalaryHike"))
# split PercentSalaryHike variable into somewhat equal-sized groups
hrdata$PercentSalaryHikeG <- as.factor(cut2(hrdata$PercentSalaryHike, g=9))
# check how each group range look like
summary(hrdata$PercentSalaryHikeG)
# group1 = 11, group2 = 12, group3 =13, group4 = 14, group5 = 15:16, group6 = 17:18, group7 = 19:21, group8=22:25
# now, convert PercentSalaryHike column(int) to nominal type
hrdata$PercentSalaryHikeG <- as.numeric(cut2(hrdata$PercentSalaryHike, g=9))
hrdata<- subset(hrdata, select = -c(PercentSalaryHike))
```

```{r}
# convert TotalWorkingYears column(int) to nominal type

# check how TotalWorkingYears's data distributed
hist(hrdata$TotalWorkingYears, breaks = 15, main = paste("Histogram of TotalWorkingYears"))
# split TotalWorkingYears variable into somewhat equal-sized groups
hrdata$TotalWorkingYearsG <- as.factor(cut2(hrdata$TotalWorkingYears, g=5))
# check how each group range look like
summary(hrdata$TotalWorkingYearsG)
# group1 = 0:5, group2 = 6:8, group3 =9:10, group4 = 11:17, group5 = 18:40
# now, convert TotalWorkingYears column(int) to nominal type
hrdata$TotalWorkingYearsG <- as.numeric(cut2(hrdata$TotalWorkingYears, g=5))
hrdata<- subset(hrdata, select = -c(TotalWorkingYears))
```

```{r}
# check how TrainingTimesLastYear's data distributed
## this could be a problem since the values are clustered between 2 to 3 but, for now keep this way.
summary(as.factor(hrdata$TrainingTimesLastYear))
```

```{r}
# convert YearsAtCompany column(int) to nominal type

# check how YearsAtCompany's data distributed
hist(hrdata$YearsAtCompany, breaks = 15, main = paste("Histogram of YearsAtCompany"))
# split YearsAtCompany variable into somewhat equal-sized groups
hrdata$YearsAtCompanyG <- as.factor(cut2(hrdata$YearsAtCompany, g=4))
# check how each group range look like
summary(hrdata$YearsAtCompanyG)
# group1 = 0:3, group2 = 4:5, group3 =6:9, group4 = 10:40
# now, convert YearsAtCompany column(int) to nominal type
hrdata$YearsAtCompanyG <- as.numeric(cut2(hrdata$YearsAtCompany, g=4))
hrdata<- subset(hrdata, select = -c(YearsAtCompany))
```

```{r}
# convert YearsSinceLastPromotion column(int) to nominal type

# check how YearsSinceLastPromotion's data distributed
hist(hrdata$YearsSinceLastPromotion, breaks = 15, main = paste("Histogram of YearsSinceLastPromotion"))
# split YearsSinceLastPromotion variable into somewhat equal-sized groups
hrdata$YearsSinceLastPromotionG <- as.factor(cut2(hrdata$YearsSinceLastPromotion, g=4))
# check how each group range look like
summary(hrdata$YearsSinceLastPromotionG)
# group1 = 0, group2 = 1, group3 =2:3, group4 = 4:15,
# now, convert YearsSinceLastPromotion column(int) to nominal type
hrdata$YearsSinceLastPromotionG <- as.numeric(cut2(hrdata$YearsSinceLastPromotion, g=4))
hrdata<- subset(hrdata, select = -c(YearsSinceLastPromotion))
```


```{r}
# convert YearsWithCurrManager column(int) to nominal type

# check how YearsWithCurrManager's data distributed
hist(hrdata$YearsWithCurrManager, breaks = 15, main = paste("Histogram of YearsWithCurrManager"))
# split YearsWithCurrManager variable into somewhat equal-sized groups
hrdata$YearsWithCurrManagerG <- as.factor(cut2(hrdata$YearsWithCurrManager, g=5))
# check how each group range look like
summary(hrdata$YearsWithCurrManagerG)
# group1 = 0:1, group2 = 2, group3 =3:4, group4 = 5:7, group5 = 8:17
# now, convert YearsWithCurrManager column(int) to nominal type
hrdata$YearsWithCurrManagerG <- as.numeric(cut2(hrdata$YearsWithCurrManager, g=5))
hrdata<- subset(hrdata, select = -c(YearsWithCurrManager))
```


```{r}
# convert Date_of_Hire column(int) to nominal type

# extract year and month from Date_of_Hire data
hrdata$Date_of_Hire <- strptime(hrdata$Date_of_Hire, "%d-%m-%Y")
hrdata$Year_of_Hire <- year(as.POSIXlt(hrdata$Date_of_Hire, format="%d/%m/%Y"))
hrdata$Month_of_Hire <- month(as.POSIXlt(hrdata$Date_of_Hire, format="%d/%m/%Y"))

# check how Year_of_Hire's data distributed
hist(hrdata$Year_of_Hire, breaks = 15, main = paste("Histogram of Year_of_Hire"))
# split Year_of_Hire variable into somewhat equal-sized groups
hrdata$Year_of_Hire_G <- as.factor(cut2(hrdata$Year_of_Hire, g=5))
# check how each group range look like
summary(hrdata$Year_of_Hire_G)
# group1 = 1969:2011, group2 = 2012:2014, group3 =2015:2016, group4 = 2017:2019, group5 = 2020:2021
# now, convert Year_of_Hire column(int) to nominal type
hrdata$Year_of_Hire_G <- as.numeric(cut2(hrdata$Year_of_Hire, g=5))
hrdata<- subset(hrdata, select = -c(Year_of_Hire))

# check how each month contains data points
summary(hrdata$Month_of_Hire)
hrdata<- subset(hrdata, select = -c(Date_of_Hire))
```


```{r}
# convert Mode_of_work column(cha) to binary type

# convert OFFICE to 1 and WFH to 0
hrdata$Mode_of_work <- ifelse(hrdata$Mode_of_work =="OFFICE",1,0)
```

```{r}
# convert Work_accident column(cha) to binary type

# convert Yes to 1 and No to 0
hrdata$Work_accident <- ifelse(hrdata$Work_accident =="Yes",1,0)
```



# correlation analysis
```{r}
hr_cor <- cor(hrdata[ ,colnames(hrdata) != "Attrition"], hrdata$Attrition)
hr_cor
```

```{r}
hist(hrdata$OverTime, breaks=2)
summary(as.factor(hrdata$OverTime))
```

```{r}
# Correlation scatterplot between Attrition and Monthly Income Group variables 
summary(as.factor(hrdata$MonthlyIncomeG))
scatterplot(Attrition~MonthlyIncomeG, smooth=TRUE, data=hrdata)
```

```{r}
# Correlation scatterplot between Attrition and Year of Hire Group variables
scatterplot(Attrition~Year_of_Hire_G, smooth=TRUE, data=hrdata)
```
```{r}
# Correlation scatterplot between Attrition and Total working years group variables
scatterplot(Attrition~TotalWorkingYearsG, smooth=TRUE, data=hrdata)
```

```{r}
# Correlation scatterplot between Attrition and Job level variables
scatterplot(Attrition~JobLevel, smooth=TRUE, data=hrdata)
```

```{r}
# Correlation scatterplot between Attrition and YearsAtCompanyG variables
scatterplot(Attrition~YearsAtCompanyG, smooth=TRUE, data=hrdata)
```

```{r}
# Correlation scatterplot between Attrition and MaritalStatus variables
scatterplot(Attrition~MaritalStatus, smooth=TRUE, data=hrdata)
```
```{r}
# Correlation scatterplot between Attrition and YearsWithCurrManagerG variables
scatterplot(Attrition~YearsWithCurrManagerG, smooth=TRUE, data=hrdata)
```

```{r}
# the elbow method
fviz_nbclust(hrdata, kmeans, method = "wss") +
  geom_vline(xintercept = 3, linetype = 2) + # add line for better visualization
  labs(subtitle = "Elbow method") # add subtitle
```

# The elbow method suggests 3 clusters

```{r}
# the silhouette method
# Silhouette method
fviz_nbclust(hrdata, kmeans, method = "silhouette") +
  labs(subtitle = "Silhouette method")
```

# The Silhouette method suggests 2 clusters.



```{r}
# generate 2 clusters
hrdata_km2 <- kmeans(hrdata, centers = 2, nstart =30)
fviz_cluster(hrdata_km2, data=hrdata)
```
```{r}
# generate 3 clusters
hrdata_km3 <- kmeans(hrdata, centers = 3, nstart =30)
fviz_cluster(hrdata_km3, data=hrdata)
```

```{r}
# generate 5 clusters
hrdata_km5 <- kmeans(hrdata, centers = 5, nstart =30)
fviz_cluster(hrdata_km5, data=hrdata)
```


```{r}
# create data frame including the 2 clusters results
hrdata_km2df <- data.frame(hrdata, hrdata_km2$cluster)

# subset of hrdata_km2df where attrition is 1 or 0
hrdata_km2at1 <- subset(hrdata_km2df, Attrition == 1)
hrdata_km2at0 <- subset(hrdata_km2df, Attrition == 0)

# randomly sampled attrition=1 or 0 data
hrdata_km2sam1 <- hrdata_km2at1[sample(1:nrow(hrdata_km2at1), size = 50, replace=FALSE), ]
hrdata_km2sam2 <- hrdata_km2at0[sample(1:nrow(hrdata_km2at0), size = 50, replace=FALSE), ]

# adds sampled data together
hrdata_km2sam <- rbind(hrdata_km2sam1,hrdata_km2sam2)

# confusion matrix to find the connection with attrition variable
table(hrdata_km2sam$Attrition,hrdata_km2sam$hrdata_km2.cluster)
```
## there is not much clear connection between attrition and cluster group

```{r}
# create data frame including the 3 clusters results
hrdata_km3df <- data.frame(hrdata, hrdata_km3$cluster)

# subset of hrdata_km3df where attrition is 1 or 0
hrdata_km3at1 <- subset(hrdata_km3df, Attrition == 1)
hrdata_km3at0 <- subset(hrdata_km3df, Attrition == 0)

# randomly sampled attrition=1 or 0 data
hrdata_km3sam1 <- hrdata_km3at1[sample(1:nrow(hrdata_km3at1), size = 50, replace=FALSE), ]
hrdata_km3sam2 <- hrdata_km3at0[sample(1:nrow(hrdata_km3at0), size = 50, replace=FALSE), ]

# adds sampled data together
hrdata_km3sam <- rbind(hrdata_km3sam1,hrdata_km3sam2)

# confusion matrix to find the connection with attrition variable
table(hrdata_km3sam$Attrition,hrdata_km3sam$hrdata_km3.cluster)
```

# Classification - Decision Tree Models

# Experimental Design
```{r}
# Now that the data is labeled, its time to design an experiment. 
# Below we randomly select a train and test set for validation

# Make Train and Test sets
numTotalem = nrow(hrdata)
trainRatio <- .60

set.seed(11) # Set Seed so that same sample can be reproduced in future also
sample <- sample.int(n = numTotalem, size = floor(trainRatio*numTotalem), replace = F)

train <- hrdata[sample, ]
test <- hrdata[-sample, ]
                                       
# train / test ratio
length(sample)/nrow(hrdata)
```


```{r}
#Train Tree Model 1
train_tree1 <- rpart(Attrition ~ ., data = train, method="class", control=rpart.control(cp=0))
summary(train_tree1)
```

```{r}
#plot the decision tree
draw.tree(train_tree1)
```

```{r}
#plot the decision tree
fancyRpartPlot(train_tree1)
```

```{r}
#predict the test dataset using the model for train tree No. 1
predicted1<- predict(train_tree1, test, type="class")
#confusion matrix to find correct and incorrect predictions
table(Attrition=predicted1, true=test$Attrition)
```


```{r}
confusionMatrix(data = predicted1, as.factor(test$Attrition))
```

# Naive Bayes Train model

```{r}
#g generate the model
train_naibayes <- naiveBayes(train$Attrition~., data=train, na.action = na.pass)
summary(train_naibayes)
```


```{r}
# Naive Bayes model Prediction
nb_Pred <- predict(train_naibayes, test)
# confusion matrix to find correct and incorrect predictions
table(Attrition=nb_Pred, true=test$Attrition)
```


# Association Rule Mining 
# data preparation
```{r}
hrdata2 <- read.csv("D:/Documents/school/IST707/project/HRdata.csv")
```

```{r}
# remove unnecessary columns
hrdata2<- subset(hrdata2, select = -c(X,Date_of_termination))
```

# convert character data type to nominal types to conduct Association Rule Mining

```{r}
# convert age column(int) to nominal type
# split age variable into somewhat equal-sized groups
hrdata2$ageG <- as.factor(cut2(hrdata2$Age, g=6))
# check how each group range look like
summary(hrdata2$ageG)
## group1 = 18:28, group2 = 29:32, group3 = 33:36, group4 = 37:40, group5 = 41:46, group6 = 47:60
hrdata2<- subset(hrdata2, select = -c(Age))
```

```{r}
# columns that are character type
chacols <- c('Attrition', 'BusinessTravel', 'Department', 'Gender', 'JobRole', 'MaritalStatus','OverTime', 'Higher_Education','Date_of_Hire', 'Status_of_leaving','Mode_of_work','Work_accident', 'Source_of_Hire', 'Job_mode')
# convert character type columns to nominal type
hrdata2[chacols] <- lapply(hrdata2[chacols], factor)
```


```{r}
# convert DistanceFromHome column(int) to nominal type
# split DistanceFromHome variable into somewhat equal-sized groups
hrdata2$DistanceFromHomeG <- as.factor(cut2(hrdata2$DistanceFromHome, g=8))
# check how each group range look like
summary(hrdata2$DistanceFromHomeG)
# group1 = 1, group2 = 2, group3 =3:4, group4 = 5:7, group5 = 8:9, group6 = 10:14, group7 = 15:22, group8=23:29
hrdata2 <- subset(hrdata2, select = -c(DistanceFromHome))
```

```{r}
# check how JobInvolvement's data distributed
hist(hrdata2$JobInvolvement, breaks = 15, main = paste("Histogram of JobInvolvement"))
# convert JobInvolvement column(int) to nominal type
# split JobInvolvement variable into somewhat equal-sized groups
hrdata2$JobInvolvement <- as.factor(cut2(hrdata2$JobInvolvement, g=4))
# check how each group range look like
summary(hrdata2$JobInvolvement)
```
```{r}
# check how JobLevel's data distributed
hist(hrdata2$JobLevel, breaks = 15, main = paste("Histogram of JobLevel"))
# convert JobLevel column(int) to nominal type
# split JobLevel variable into somewhat equal-sized groups
hrdata2$JobLevel <- as.factor(cut2(hrdata2$JobLevel, g=3))
# check how each group range look like
summary(hrdata2$JobLevel)
```
```{r}
# check how JobSatisfaction's data distributed
hist(hrdata2$JobSatisfaction, breaks = 15, main = paste("Histogram of JobSatisfaction"))
# convert JobSatisfaction column(int) to nominal type
# split JobSatisfaction variable into somewhat equal-sized groups
hrdata2$JobSatisfaction <- as.factor(cut2(hrdata2$JobSatisfaction, g=3))
# check how each group range look like
summary(hrdata2$JobSatisfaction)
```

```{r}
# convert MonthlyIncome column(int) to nominal type

# split MonthlyIncome variable into somewhat equal-sized groups
hrdata2$MonthlyIncomeG <- as.factor(cut2(hrdata2$MonthlyIncome, g=6))
# check how each group range look like
summary(hrdata2$MonthlyIncomeG)
# group1 = 1009:2560, group2 = 2561:3632, group3 =3633:4929, group4 = 4930:6537, group5 = 6538:10551, group6 = 10552:19999
hrdata2<- subset(hrdata2, select = -c(MonthlyIncome))
```

```{r}
# check how NumCompaniesWorked's data distributed
hist(hrdata2$NumCompaniesWorked, breaks = 15, main = paste("Histogram of NumCompaniesWorked"))
# split MonthlyIncome variable into somewhat equal-sized groups
hrdata2$NumCompaniesWorked <- as.factor(cut2(hrdata2$NumCompaniesWorked, g=8))
# check how each group range look like
summary(hrdata2$NumCompaniesWorked)
```
```{r}
# check the OverTime variable value
summary(hrdata2$OverTime)
```

```{r}
# convert PercentSalaryHike column(int) to nominal type

# check how PercentSalaryHike's data distributed
hist(hrdata2$PercentSalaryHike, breaks = 15, main = paste("Histogram of PercentSalaryHike"))
# split PercentSalaryHike variable into somewhat equal-sized groups
hrdata2$PercentSalaryHike <- as.factor(cut2(hrdata2$PercentSalaryHike, g=9))
# check how each group range look like
summary(hrdata2$PercentSalaryHike)
```
```{r}
# convert PerformanceRating column(int) to nominal type
# split PercentSalaryHike variable into somewhat equal-sized groups
hrdata2$PerformanceRating <- as.factor(hrdata2$PerformanceRating)
# check the OverTime variable value
summary(hrdata2$PerformanceRating)
```
```{r}
# convert StockOptionLevel column(int) to nominal type
# split StockOptionLevel variable into somewhat equal-sized groups
hrdata2$StockOptionLevel <- as.factor(cut2(hrdata2$StockOptionLevel, g=3))
# check how each group range look like
summary(hrdata2$StockOptionLevel)
```

```{r}
# convert TotalWorkingYears column(int) to nominal type

# check how TotalWorkingYears's data distributed
hist(hrdata2$TotalWorkingYears, breaks = 15, main = paste("Histogram of TotalWorkingYears"))
# split TotalWorkingYears variable into somewhat equal-sized groups
hrdata2$TotalWorkingYears <- as.factor(cut2(hrdata2$TotalWorkingYears, g=5))
# check how each group range look like
summary(hrdata2$TotalWorkingYears)
```


```{r}
# check how TrainingTimesLastYear's data distributed
# check how TrainingTimesLastYear's data distributed
hist(hrdata2$TrainingTimesLastYear, breaks = 15, main = paste("Histogram of TrainingTimesLastYear"))
# split TrainingTimesLastYear variable into somewhat equal-sized groups
hrdata2$TrainingTimesLastYear <- as.factor(cut2(hrdata2$TrainingTimesLastYear, g=4))
# check how each group range look like
summary(hrdata2$TrainingTimesLastYear)

```

```{r}
# convert YearsAtCompany column(int) to nominal type

# check how YearsAtCompany's data distributed
hist(hrdata2$YearsAtCompany, breaks = 15, main = paste("Histogram of YearsAtCompany"))
# split YearsAtCompany variable into somewhat equal-sized groups
hrdata2$YearsAtCompany <- as.factor(cut2(hrdata2$YearsAtCompany, g=4))
# check how each group range look like
summary(hrdata2$YearsAtCompany)
```

```{r}
# convert YearsSinceLastPromotion column(int) to nominal type

# check how YearsSinceLastPromotion's data distributed
hist(hrdata2$YearsSinceLastPromotion, breaks = 15, main = paste("Histogram of YearsSinceLastPromotion"))
# split YearsSinceLastPromotion variable into somewhat equal-sized groups
hrdata2$YearsSinceLastPromotion <- as.factor(cut2(hrdata2$YearsSinceLastPromotion, g=4))
# check how each group range look like
summary(hrdata2$YearsSinceLastPromotion)
```


```{r}
# convert YearsWithCurrManager column(int) to nominal type

# check how YearsWithCurrManager's data distributed
hist(hrdata2$YearsWithCurrManager, breaks = 15, main = paste("Histogram of YearsWithCurrManager"))
# split YearsWithCurrManager variable into somewhat equal-sized groups
hrdata2$YearsWithCurrManager <- as.factor(cut2(hrdata2$YearsWithCurrManager, g=5))
# check how each group range look like
summary(hrdata2$YearsWithCurrManager)
```


```{r}
# convert Date_of_Hire column(int) to nominal type

# extract year and month from Date_of_Hire data
hrdata2$Date_of_Hire <- strptime(hrdata2$Date_of_Hire, "%d-%m-%Y")
hrdata2$Year_of_Hire <- year(as.POSIXlt(hrdata2$Date_of_Hire, format="%d/%m/%Y"))
hrdata2$Month_of_Hire <- month(as.POSIXlt(hrdata2$Date_of_Hire, format="%d/%m/%Y"))

# check how Year_of_Hire's data distributed
hist(hrdata2$Year_of_Hire, breaks = 15, main = paste("Histogram of Year_of_Hire"))
# split Year_of_Hire variable into somewhat equal-sized groups
hrdata2$Year_of_Hire_G <- as.factor(cut2(hrdata2$Year_of_Hire, g=5))
# check how each group range look like
summary(hrdata2$Year_of_Hire_G)

hrdata2<- subset(hrdata2, select = -c(Year_of_Hire))

# check how each month contains data points
hrdata2$Month_of_Hire <- as.factor(hrdata2$Month_of_Hire)
summary(hrdata2$Month_of_Hire)
hrdata2<- subset(hrdata2, select = -c(Date_of_Hire))
```


```{r}
# convert Leaves column(int) to nominal type

# check how Leaves's data distributed
hist(hrdata2$Leaves, breaks = 15, main = paste("Histogram of Leaves"))
# evenly distributed, thus dose need not be split
hrdata2$Leaves <- as.factor(hrdata2$Leaves)
```
```{r}
# convert Absenteeism column(int) to nominal type

# check how Absenteeism's data distributed
hist(hrdata2$Absenteeism, breaks = 15, main = paste("Histogram of Absenteeism"))
# evenly distributed, thus dose need not be split
hrdata2$Absenteeism <- as.factor(hrdata2$Absenteeism)
```

```{r}
# the most unevenly distributed variables
# OverTime and PerformanceRating
summary(hrdata2$OverTime)
summary(hrdata2$PerformanceRating)
```

```{r}
# subset the data frame
# majority(OverTime=No and PerformanceRating=3) group
hrdatat2ot0pf3 <- hrdata2[(hrdata2$OverTime == "No") & (hrdata2$PerformanceRating == 3),]
# minority(OverTime=Yes and PerformanceRating=4) group
hrdatat2ot1pf4 <- hrdata2[(hrdata2$OverTime == "Yes") & (hrdata2$PerformanceRating == 4),]

set.seed(1234)
# randomly sampled majority and minority group dataset
hrdata_maj <- hrdatat2ot0pf3[sample(1:nrow(hrdatat2ot0pf3), size = 50, replace=FALSE), ]
hrdata_min <- hrdatat2ot1pf4[sample(1:nrow(hrdatat2ot1pf4), size = 50, replace=FALSE), ]

# combined sampled data together
hrdata_com <- rbind(hrdata_maj,hrdata_min)

# OverTime and PerformanceRating
summary(hrdata_com$OverTime)
summary(hrdata_com$PerformanceRating)
```

# conducted Association Rule Mining

```{r}
# test rules: targeting employees who stayed
hr_rules1<- apriori(hrdata_com, parameter = list(support = 0.07, confidence = 1), appearance = list(default="lhs", rhs="Attrition=No"), control=list(verbose=F))
summary(hr_rules1)
```

```{r}
# sort the same set of rules to view based on support
hr_rulessupport<- sort(hr_rules1, by="support", decreasing=TRUE)
inspect(hr_rulessupport[1:20])
```

```{r}
#plot the rules
hr_rulessupport <- hr_rulessupport[1:10]
plot(hr_rulessupport, method="graph")
```

```{r}
# sort the same set of rules to view based on lift
hr_ruleslift<- sort(hr_rules1, by="lift", decreasing=TRUE)
inspect(hr_ruleslift[1:20])
```


```{r}
summary(hr_ruleslift)
#plot the rules
hr_rulesliftbest <- hr_ruleslift[1:10]
plot(hr_rulesliftbest, method="graph")
```




